---
title: "p8105_hw2_xz2981"
author: "Xinyan Zhou"
date: "2021/10/4"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


# Problem 1 

## Read and clean the Mr. Trash Wheel sheet

```{r warning=FALSE}
p1_a <- 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
            sheet = "Mr. Trash Wheel", range = "A2:N535") %>%
  janitor::clean_names() %>%
  filter(dumpster > 0) %>%
  mutate(sports_balls = round(sports_balls, digits = 0))

p1_a_2019 <-  filter(p1_a,year == 2019)
```

### Data description
The dataset contains `r nrow(p1_a)` observations of `r ncol(p1_a)` variables, and key variables include dumpster, month, year, date, 
weight_tons,and homes_powered.
 
The median number of sports balls in a dumpster in 2019 is 
`r round(median(pull(p1_a_2019,sports_balls), na.rm = TRUE))`.

## Read and clean precipitation data for 2018 and 2019

```{r}
p1_b_2018 <-
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2018 Precipitation",range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2018)
```

```{r}
p1_b_2019 <-
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2019 Precipitation",range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2019)
```

```{r}
p1_b_all <-
  bind_rows(p1_b_2018,p1_b_2019) %>%
  mutate(month = month.name[month])
```

### Data description
The dataset contains `r nrow(p1_b_all)` observations of `r ncol(p1_b_all)` variables, and key variables include month, total, year.
 
The total precipitation in 2018 is `r (sum(pull(p1_b_2018,total)))`.

# Problem 2

## Clean the data in pols-month.csv

```{r}
p2_a <-
  read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon,c("year","month","day"),sep = "-",convert = TRUE) %>%
  mutate(month = month.name[month], president = "0") %>%
  mutate(president = replace(president, prez_gop == 1, "gop")) %>%
  mutate(president = replace(president, prez_dem == 1, "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
```

## Clean the data in snp.csv

```{r}
p2_b <-
  read_csv(file = "./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date,c("month","day","year"),sep = "/",convert = TRUE) %>%
  mutate(month = month.name[month]) %>%
  mutate(year = if_else(year > 20, year + 1900, year + 2000)) %>%
  arrange(year,month) %>% 
  select(year,month,close)
```

## Tidy the unemployment data

```{r}
p2_c <-
  read_csv(file = "./data/unemployment.csv") %>%
  rename("1" = Jan,"2" = Feb, "3" = Mar, "4" =  Apr, "5" = May, 
         "6" = Jun, "7" = Jul, "8" = Aug, "9" = Sep, "10" = Oct,
         "11" = Nov, "12" = Dec) %>%
  pivot_longer(
    "1":"12",
    names_to = "month",
    values_to = "unemployment") %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.name[month]) %>%
  janitor::clean_names()
```

## Join them!
```{r}
p2_d <-
  left_join(p2_a,p2_b,by = c("year","month"))
p2_final <-
  left_join(p2_d,p2_c,by = c("year","month"))

```

## Descriptions for each dataset
p2_a (based on pols-month.csv): The dataset contains `r nrow(p2_a)` observations of `r ncol(p2_a)` variables, and key variables include year, month, and president.

p2_b (based on snp.csv): The dataset contains `r nrow(p2_b)` observations of `r ncol(p2_b)` variables, and key variables include year, month, and close.

p2_c (based on unemployment.csv): The dataset contains `r nrow(p2_c)` observations of `r ncol(p2_c)` variables, and key variables include year, month, and unemployment.

## Descriptions for the resulting dataset
The resulting dataset contains `r nrow(p2_final)` observations of `r ncol(p2_final)` variables. The range of year is from `r min(pull(p2_final,year))` to `r max(pull(p2_final,year))`. 
Key variables include year, month, president, close, and employment.


# Problem 3

```{r}
p3_a <-
  read_csv(file = "./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = str_to_title(ethnicity),
         gender = str_to_title(gender),
         childs_first_name = str_to_title(childs_first_name)) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "Asian And Paci",
                             "Asian And Pacific Islander")) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "Black Non Hisp",
                             "Black Non Hispanic")) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "White Non Hisp",
                             "White Non Hispanic")) %>%
  distinct()
```

## How the rank in popularity of the name “Olivia” as a female baby name changes over time

```{r}
p3_Olivia <-
  filter(p3_a, gender == "Female", childs_first_name == "Olivia") %>%
  select(year_of_birth, ethnicity, rank)

table_Olivia <-
  pivot_wider(p3_Olivia, names_from = "year_of_birth", values_from = "rank")

table_Olivia
```

## The most popular name among male children over time

```{r}
p3_male <-
  filter(p3_a, gender == "Male", rank == 1) %>%
  select(year_of_birth, ethnicity, childs_first_name)

table_male <-
  pivot_wider(p3_male, names_from = "year_of_birth", 
        values_from = "childs_first_name")

table_male
```

## Plot

```{r}
p3_plot <-
  filter(p3_a, gender == "Male", year_of_birth == 2016, 
         ethnicity == "White Non Hispanic")

plot <-
  ggplot(p3_plot, aes(rank,count)) + geom_point()
plot
```



